\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,booktabs,hyperref}
\title{IMU Error Model}
\author{imu-error-model package}
\date{\today}

\begin{document}
\maketitle

\section{Purpose and boundary}
The package models the measurement errors of a three-axis accelerometer and
gyroscope. It consumes truth velocity with gravity excluded and truth
orientation. It does not implement rigid-body dynamics, gravity, navigation,
or state reconstruction.

\section{Truth-state contract}
Let $R^w_b(t)$ map body-frame vectors into the world frame and let
$v^w(t)$ be truth velocity with gravity excluded. For an interval beginning at
$t_0$ and ending at $t_1$, the ideal increments are
\begin{align}
\Delta v_b &= (R^w_b(t_0))^T\left[v^w(t_1)-v^w(t_0)\right],\\
\Delta\theta_b &= \operatorname{Log}\left((R^w_b(t_0))^T R^w_b(t_1)\right)^\vee.
\end{align}
Both are expressed in the body frame at $t_0$.

The public \texttt{measure} contract takes a timestamp in seconds, a
three-vector $v^w$ in m/s, a proper $3\times3$ rotation matrix, and an
optional temperature in degrees Celsius. The first call establishes the
initial state and returns zero increments; later calls require strictly
increasing timestamps. \texttt{ImuOutput.delta\_v} is in m/s and
\texttt{ImuOutput.delta\_theta} is in radians.

\section{Implemented measurement equation}
Let
\begin{equation}
u_{a,k} = \frac{\Delta v_{b,k}}{\Delta t_k}, \qquad
u_{g,k} = \frac{\Delta\theta_{b,k}}{\Delta t_k}
\end{equation}
be the ideal accelerometer and gyroscope rates for one sample interval. For
sensor $s\in\{a,g\}$, define the temperature offset and the three linear
thermal terms as
\begin{equation}
\delta T_k = T_k-T_{0,s}, \qquad
\beta_{T,s} = c_{b,s}\delta T_k, \qquad
\lambda_{T,s} = 1+c_{n,s}\delta T_k, \qquad
\kappa_{T,s}=c_{\mathrm{sf},s}\delta T_k.
\end{equation}
The coefficients may be scalars or three-vectors; vector products below are
elementwise. The default thermal model requires
$\lambda_{T,s}\geq0$ componentwise.

The deterministic scale and nonlinearity stage is
\begin{equation}
d_s(u_{s,k},T_k) =
u_{s,k}\odot
\left(1+s_s\mathbf{1}+\kappa_{T,s}\right)
\odot\left(1+n_s\lVert u_{s,k}\rVert_2^2\right)
+\beta_{T,s},
\end{equation}
where $s_s$ is the fixed scale-factor coefficient and $n_s$ is the
nonlinearity coefficient. A turn-on misalignment vector
$\epsilon_s\sim\mathcal{N}(0,m_s^2 I)$ is sampled at reset and held fixed for
the run. Its rotation matrix is
\begin{equation}
M_s=\exp\left([\epsilon_s]_\times\right).
\end{equation}

The pre-output rate is then
\begin{equation}
r_{s,k}=M_s d_s(u_{s,k},T_k)
+b_{\mathrm{on},s}+b_{s,k}+f_{s,k}+n_{s,k}.
\end{equation}
Clipping, when enabled, and quantization are applied to this rate:
\begin{align}
\bar r_{s,k} &=
\operatorname{clip}(r_{s,k},-L_s,L_s),\\
q_{s,k} &=
\begin{cases}
\delta_s\operatorname{round}(\bar r_{s,k}/\delta_s),
&\text{if a quantization step }\delta_s\text{ is configured},\\
\bar r_{s,k},&\text{otherwise}.
\end{cases}
\end{align}
The reported increments include both per-channel and global output scales:
\begin{equation}
\Delta v^{\mathrm{meas}}_k =
\Delta t_k\,o_a O_a q_{a,k}, \qquad
\Delta\theta^{\mathrm{meas}}_k =
\Delta t_k\,o_g O_g q_{g,k}.
\end{equation}
Here $L_s$ is the measurement range, $o_s$ is the channel output scale, and
$O_s$ is the corresponding \texttt{ImuConfig} output scale. If clipping is
disabled, $\operatorname{clip}$ is the identity.

\section{Stochastic processes}
White rate noise with spectral density $N$ is sampled with standard deviation
$N/\sqrt{\Delta t}$. A stationary first-order Gauss--Markov bias is initialized
with $b_0\sim\mathcal{N}(0,\sigma_b^2I)$ and updated as
\begin{equation}
b_{k+1}=\phi b_k+\eta_k, \qquad
\phi=e^{-\Delta t/T}, \qquad
\eta_k\sim\mathcal{N}(0,\sigma_b^2(1-\phi^2)I).
\end{equation}
When no correlation time is supplied, the same parameter is interpreted as a
random-walk driving standard deviation and the update is instead
\begin{equation}
b_{k+1}=b_k+\eta_k, \qquad
\eta_k\sim\mathcal{N}(0,\sigma_b^2\Delta t\,I).
\end{equation}
In both cases, the turn-on bias is independently initialized as
$b_{\mathrm{on}}\sim\mathcal{N}(0,\sigma_{\mathrm{on}}^2I)$ and remains fixed
until the next reset.

For a supplied continuous-time covariance $C_s$, define
\begin{equation}
\Lambda_{s,k}=
\begin{cases}
\lambda_{T,s}I,&\lambda_{T,s}\text{ is scalar},\\
\operatorname{diag}(\lambda_{T,s}),&\lambda_{T,s}\text{ is a vector}.
\end{cases}
\end{equation}
The white rate noise used by the model is
\begin{equation}
n_{s,k}\sim
\begin{cases}
\mathcal{N}\left(0,\dfrac{N_s^2\Lambda_{s,k}\Lambda_{s,k}^{T}}{\Delta t_k}\right),
& C_s\text{ is not supplied},\\[8pt]
\mathcal{N}\left(0,\dfrac{\Lambda_{s,k}C_s\Lambda_{s,k}^{T}}{\Delta t_k}\right),
& C_s\text{ is supplied}.
\end{cases}
\end{equation}
When $C_s$ is supplied, it replaces the independent-axis density sampling
path.

\subsection{Allan-deviation interpretation}
The analysis scripts use non-overlapping blocks of $m$ rate samples. With
$J$ complete block means, the averaging time and estimator are
\begin{align}
\tau &= m\Delta t, \qquad
\bar y_j^{(m)}=\frac{1}{m}\sum_{i=jm}^{(j+1)m-1}y_i,\\
\widehat{\sigma}_A^2(\tau)
&=\frac{1}{2(J-1)}
\sum_{j=0}^{J-2}\left(\bar y_{j+1}^{(m)}-\bar y_j^{(m)}\right)^2.
\end{align}
For independent white rate noise, this estimator has the reference behavior
\begin{equation}
\sigma_A(\tau)=\frac{N}{\sqrt{\tau}}
\end{equation}
at the valid discrete averaging times $\tau=m\Delta t$.
This is consistent with the sample-level standard deviation
$N/\sqrt{\Delta t}$ used by the model and with the estimator used in
\texttt{examples/allan\_variance.py}.

For a stationary Gauss--Markov process with stationary standard deviation
$\sigma_b$ and correlation time $T$, the exact continuous-time reference
curve is
\begin{equation}
\sigma_A^2(\tau) =
\sigma_b^2\left[
\frac{2T}{\tau}
- \left(\frac{T}{\tau}\right)^2
\left(3 - 4e^{-\tau/T} + e^{-2\tau/T}\right)
\right].
\end{equation}
It has an approximately $+1/2$ slope at $\tau \ll T$, turns near $T$, and
has an approximately $-1/2$ slope at $\tau \gg T$. If $T$ is omitted, the
implementation uses a random walk instead: \texttt{bias\_std} is then the
driving scale and the long-time Allan deviation has an approximately $+1/2$
slope rather than a stationary turnover.

An optional \texttt{noise\_covariance} is a positive-semidefinite $3\times3$
continuous-time covariance for the corresponding rate noise. It is used in
place of independent-axis white-noise sampling. Accelerometer densities are
in m/s$^2/\sqrt{\mathrm{Hz}}$ and gyro densities are in
rad/s$/\sqrt{\mathrm{Hz}}$.

The turn-on bias is sampled once at reset and remains fixed for that simulated
run. It is intentionally separate from the in-run Gauss--Markov state; this
prevents a repeatability error from being mistaken for a time-correlated drift
process in Allan analysis.

To approximate a finite bias-instability plateau, the package can sum several
independent Gauss--Markov processes with logarithmically spaced correlation
times:
\begin{equation}
b_f(t) = \sum_{i=1}^{K} x_i(t), \qquad
T_i \in \operatorname{geomspace}(T_{\min},T_{\max},K).
\end{equation}
Each component is initialized with
$x_{i,0}\sim\mathcal{N}(0,s_i^2I)$ and advanced exactly as
\begin{equation}
x_{i,k+1}=\phi_i x_{i,k}+\eta_{i,k}, \qquad
\phi_i=e^{-\Delta t/T_i}, \qquad
\eta_{i,k}\sim\mathcal{N}(0,s_i^2(1-\phi_i^2)I).
\end{equation}
Let $\mathcal{A}(v,T,\tau)$ denote the Allan variance of a stationary
Gauss--Markov component with stationary variance $v$:
\begin{equation}
\mathcal{A}(v,T,\tau)=
\frac{2v\left(T\tau-T^2(1-e^{-\tau/T})\right)
-vT^2(1-e^{-\tau/T})^2}{\tau^2}.
\end{equation}
At the geometric-center averaging time
\begin{equation}
\tau_c = \sqrt{T_{\min}T_{\max}}.
\end{equation}
the implementation uses equal component standard deviations
\begin{equation}
s_i =
\frac{\sigma_f}
{\sqrt{\displaystyle\sum_{\ell=1}^{K}\mathcal{A}(1,T_\ell,\tau_c)}},
\qquad
\sigma_f=\texttt{flicker\_bias\_std}.
\end{equation}
Thus \texttt{flicker\_bias\_std} is an Allan-deviation target for the sum, not
the standard deviation of one component. This is a finite-band $1/f$-like
approximation: it produces a plateau-like region and rolls off outside the
configured correlation-time band. Increasing \texttt{flicker\_components}
makes the approximation smoother but does not widen that band. The process is
not intended to represent an infinite-band stationary $1/f$ process.

For independent stochastic terms, Allan variances add. Finite records and
random seeds still produce Monte Carlo scatter around the reference curves, so
the parity tests compare the white-noise level and flicker target within
tolerances rather than requiring every simulated point to lie exactly on an
analytic curve. The short-correlation and finite-band fixtures under
\texttt{profiles/test/} provide compact regression cases.

\section{Extension and verification}
Alternate implementations satisfy \texttt{ImuModelProtocol} or inherit
\texttt{BaseImuModel}. The repository's Python task runner provides local and
GitHub-compatible test, coverage, lint, package-build, and documentation
commands. Notional hardware-estimate YAML profiles are retained under the
examples tree. Every numeric value in those profiles is approximate and
non-official, including values anchored to datasheets; they are not vendor
specifications, certification results, or performance guarantees. The optional
YAML adapter maps their fields into the current schema without rewriting the
source files and preserves their metadata and reference links.

The package is distributed as \texttt{imu-error-model} under the MIT License.
The core package does not include vendor-certified configurations; the
hardware-oriented profiles are illustrative research examples only.

\end{document}
